---
title: "Data621_Final_Project"
author: "Deepak Mongia"
date: "5/15/2020"
output:
  html_document: default
  pdf_document: default
---

### Data 621 - Final Project

For this project, we are going to work on a Housing dataset which gives around 80 independent features of around 2900 houses and their locality in the city of Ames, Iowa. The dataset was actually created for academia by a professor using the raw data coming from the Assessor’s Office in the form of a data dump from their records system.
As a part of this project, we are analyzing this raw data and understand how the sale price is dependent on these features.

#### Data - Introduction
Ask a home buyer to describe their dream house, and they probably won't begin with the height of the basement ceiling or the proximity to an east-west railroad. But this playground competition's dataset proves that much more influences price negotiations than the number of bedrooms or a white-picket fence.
With 79 explanatory variables describing (almost) every aspect of residential homes in Ames, Iowa, this is a real world data set which needs to be explored further to understand how the Sale Price of a house in the city of Ames, Iowa can be determined.

Data Source: https://www.kaggle.com/c/house-prices-advanced-regression-techniques/data

Out of the 80 total variables, 44 are categorical variables like: Street: Type of road access to property which has 2 values - "Grvl" (for Gravel) and "Pave" for Paved. Remaining 36 variables are numeric features.
Response or Target feature is the Sale price of a unit which is going to be our element / feature of interest for the prediction - based on the independent features.

All the variables description is given in this file:
https://raw.githubusercontent.com/deepakmongia/Data621/master/Final%20Project/Data/data_description.txt

Let us now start with the work.
We will load our required libraries first, load the dataset, and see the basic EDA features of the dataset.

###### Loading the libraries
```{r}
library(ggplot2)
library(ggmap)
library(cowplot)
library(corrplot)
library(dplyr)
library(mice)
library(VIM)
library(corrplot)
library(gridExtra)
library(ggResidpanel)
library(pdp)

```

###### Loading the dataset and exploring the dataset
```{r}
housing2_raw <- read.csv("https://raw.githubusercontent.com/deepakmongia/Data621/master/Final%20Project/Data/train.csv",
                        header = TRUE, row.names = 1)


dim(housing2_raw)

colnames(housing2_raw)

summary(housing2_raw)

head(housing2_raw)

```

###### Check if we have any missing data
Let us first check if we have any columns with blank data
```{r}
train_missing_df1 <- data.frame(apply(housing2_raw, 2, function(x) length(which(x == ''))))
colnames(train_missing_df1) <- "counts"
train_missing_df1$column_name <- rownames(train_missing_df1)
train_missing_df1 <- data.frame(train_missing_df1[train_missing_df1$counts > 0,])
print(train_missing_df1)

```

So there is no data which is blank.

Let us check NA data now
```{r}
train_na_df1 <- data.frame(apply(housing2_raw, 2, function(x) length(which(is.na(x)))))

colnames(train_na_df1) <- "counts"
train_na_df1$column_name <- rownames(train_na_df1)
train_na_df1 <- data.frame(train_na_df1[train_na_df1$counts > 0,])
print(train_na_df1)

```

So, we have NA data in both numeric as well as categorical data.
We will deal with NA data ahead.

###### Impute numeric data
Let us now impute the missing / NA numeric features
```{r}
#### Impute Numeric data
for (i in nrow(housing2_raw)) {
  housing2_raw$LotFrontage <- ifelse(is.na(housing2_raw$LotFrontage), 
                                     median(housing2_raw$LotFrontage, na.rm = TRUE),
                                     housing2_raw$LotFrontage)
}

for (i in nrow(housing2_raw)) {
  housing2_raw$MasVnrArea <- ifelse(is.na(housing2_raw$MasVnrArea), 
                                     median(housing2_raw$MasVnrArea, na.rm = TRUE),
                                     housing2_raw$MasVnrArea)
}

head(housing2_raw$GarageYrBlt,50)
median(housing2_raw$GarageYrBlt, na.rm = TRUE)

for (i in nrow(housing2_raw)) {
  housing2_raw$GarageYrBlt <- ifelse(is.na(housing2_raw$GarageYrBlt), 
                                     median(housing2_raw$GarageYrBlt, na.rm = TRUE),
                                     housing2_raw$GarageYrBlt)
}

housing2_raw$MoSold <- as.factor(housing2_raw$MoSold)

```

###### Plot Numeric data
Let us plot the numeric data now.
```{r}
numeric_housing2_raw <- Filter(is.numeric, housing2_raw)

hist_plots_list <- list()
for (i in 1:length(colnames(numeric_housing2_raw))){
  print(i)
  p1 <- eval(substitute(
    ggplot(numeric_housing2_raw, aes(numeric_housing2_raw[, i])) + geom_histogram() +
      xlab(colnames(numeric_housing2_raw)[i])
  ,list(i = i)))
  
  hist_plots_list[[i]] <- p1
}

```

```{r}
hist_plots_list[[1]]
hist_plots_list[[2]]
hist_plots_list[[3]]
hist_plots_list[[4]]
hist_plots_list[[5]]
hist_plots_list[[6]]
hist_plots_list[[7]]
hist_plots_list[[8]]
hist_plots_list[[9]]
hist_plots_list[[10]]
hist_plots_list[[11]]
hist_plots_list[[12]]
hist_plots_list[[13]]
hist_plots_list[[14]]
hist_plots_list[[15]]
hist_plots_list[[16]]
hist_plots_list[[17]]
hist_plots_list[[18]]
hist_plots_list[[19]]
hist_plots_list[[20]]
hist_plots_list[[21]]
hist_plots_list[[22]]
hist_plots_list[[23]]
hist_plots_list[[24]]
hist_plots_list[[25]]
hist_plots_list[[26]]
hist_plots_list[[27]]
hist_plots_list[[28]]
hist_plots_list[[29]]
hist_plots_list[[30]]
hist_plots_list[[31]]
hist_plots_list[[32]]
hist_plots_list[[33]]
hist_plots_list[[34]]
hist_plots_list[[35]]
hist_plots_list[[36]]


```

###### Factor data plots
Plotting the plots for the factors now - to see how levels for each categorical feature impact the SalePrice
```{r}
factor_housing2_raw <- Filter(is.factor, housing2_raw)
factor_housing2_raw <- cbind(factor_housing2_raw, housing2_raw$SalePrice)

dim(factor_housing2_raw)

colnames(factor_housing2_raw)[colnames(factor_housing2_raw) == 'housing2_raw$SalePrice'] <-
  "SalePrice"

box_plots_list <- list()
for (i in 1:(length(colnames(factor_housing2_raw)) - 1)){
  print(i)
  p1 <- eval(substitute(
    ggplot(factor_housing2_raw, aes(x = factor_housing2_raw[, i], y = SalePrice)) + 
      geom_boxplot() +
      xlab(colnames(factor_housing2_raw)[i])
    ,list(i = i)))
  
  box_plots_list[[i]] <- p1
}

```

```{r}
box_plots_list[[1]]
box_plots_list[[2]]
box_plots_list[[3]]
box_plots_list[[4]]
box_plots_list[[5]]
box_plots_list[[6]]
box_plots_list[[7]]
box_plots_list[[8]]
box_plots_list[[9]]
box_plots_list[[10]]
box_plots_list[[11]]
box_plots_list[[12]]
box_plots_list[[13]]
box_plots_list[[14]]
box_plots_list[[15]]
box_plots_list[[16]]
box_plots_list[[17]]
box_plots_list[[18]]
box_plots_list[[19]]
box_plots_list[[20]]
box_plots_list[[21]]
box_plots_list[[22]]
box_plots_list[[23]]
box_plots_list[[24]]
box_plots_list[[25]]
box_plots_list[[26]]
box_plots_list[[27]]
box_plots_list[[28]]
box_plots_list[[29]]
box_plots_list[[30]]
box_plots_list[[31]]
box_plots_list[[32]]
box_plots_list[[33]]
box_plots_list[[34]]
box_plots_list[[35]]
box_plots_list[[36]]
box_plots_list[[37]]
box_plots_list[[38]]
box_plots_list[[39]]
box_plots_list[[40]]
box_plots_list[[41]]
box_plots_list[[42]]
box_plots_list[[43]]
box_plots_list[[44]]

```

##### Correlation
We will check now the correlation among the numerical variables
```{r}
#### Correlation

corrMatrix <- round(cor(numeric_housing2_raw),4)

corrMatrix %>% corrplot(., method = "color", outline = T, addgrid.col = "darkgray", 
                        order="hclust", addrect = 4, rect.col = "black", 
                        rect.lwd = 5,cl.pos = "b", tl.col = "indianred4", 
                        tl.cex = 1.0, cl.cex = 1.0, addCoef.col = "white", 
                        number.digits = 2, number.cex = 0.8, 
                        col =
                          colorRampPalette(c("darkred","white","dodgerblue4"))(100))

cor_numVar <- cor(numeric_housing2_raw, use="pairwise.complete.obs") #correlations of all numeric variables

#sort on decreasing correlations with SalePrice
cor_sorted <- as.matrix(sort(cor_numVar[,'SalePrice'], decreasing = TRUE))
#select only high corelations
CorHigh <- names(which(apply(cor_sorted, 1, function(x) abs(x)>0.5)))
cor_numVar <- cor_numVar[CorHigh, CorHigh]

corrplot.mixed(cor_numVar, tl.col="black", tl.pos = "lt")

```

Checking the relationship between OverallQual and SalePrice
```{r}
### SalePrice as compared with OverallQual

ggplot(data = numeric_housing2_raw, aes(x = OverallQual, y = SalePrice)) +
  geom_point()

ggplot(data = numeric_housing2_raw, aes(x = as.factor(OverallQual), y = SalePrice)) +
  geom_boxplot()

### SalePrice as compared with GrLivArea

ggplot(data = numeric_housing2_raw, aes(x = GrLivArea, y = SalePrice)) +
  geom_point()

### SalePrice as compared with GarageCars

ggplot(data = numeric_housing2_raw, aes(x = as.factor(GarageCars), y = SalePrice)) +
  geom_boxplot()

```

These plots show clear relationship of SalePrice with these 3 numeric features:
Overall Quality - OverallQual
GrLivArea - Above grade (ground) living area square feet
GarageCars - Garage location


#### Recoding the categorical features
We will now work on recoding the categorical features so that they are ready to be fed into the regression models.

```{r}
##MSZoning
summary(housing2_raw$MSZoning)
housing2_raw$Zone[housing2_raw$MSZoning == "FV"] <- 4
housing2_raw$Zone[housing2_raw$MSZoning == "RL"] <- 3
housing2_raw$Zone[housing2_raw$MSZoning == "RH" | housing2_raw$MSZoning == "RM"] <- 2
housing2_raw$Zone[housing2_raw$MSZoning == "C (all)"] <- 1

table(housing2_raw$Zone)
housing2_raw$MSZoning <- NULL

##Street
summary(housing2_raw$Street)

housing2_raw$street_paved[housing2_raw$Street == "Pave"] <- 1
housing2_raw$street_paved[housing2_raw$Street == "Grvl"] <- 0

table(housing2_raw$street_paved)
housing2_raw$Street <- NULL

##3 Alley
summary(housing2_raw$Alley)

levels(housing2_raw$Alley) <- c(levels(housing2_raw$Alley), "None")

housing2_raw$Alley[is.na(housing2_raw$Alley)] <- "None"

housing2_raw$Paved_alley[housing2_raw$Alley == "Pave"] <- 1
housing2_raw$Paved_alley[!housing2_raw$Alley == "Pave"] <- 0

table(housing2_raw$Paved_alley)
housing2_raw$Alley <- NULL

##4 LotShape
summary(housing2_raw$LotShape)

housing2_raw$Lot_regular[housing2_raw$LotShape == "Reg"] <- 1
housing2_raw$Lot_regular[!housing2_raw$LotShape == "Reg"] <- 0

table(housing2_raw$Lot_regular)
housing2_raw$LotShape <- NULL

##5 LandContour
summary(housing2_raw$LandContour)

housing2_raw$LandContour_level[housing2_raw$LandContour == "Lvl"] <- 1
housing2_raw$LandContour_level[!housing2_raw$LandContour == "Lvl"] <- 0

table(housing2_raw$LandContour_level)
housing2_raw$LandContour <- NULL

##6 Utilities
summary(housing2_raw$Utilities)

### We can drop this column
housing2_raw$Utilities <- NULL

##7 LotConfig
summary(housing2_raw$LotConfig)

housing2_raw$LotConfig_culdsac_fr3[housing2_raw$LotConfig == "CulDSac" | housing2_raw$LotConfig == "FR3"] <- 1
housing2_raw$LotConfig_culdsac_fr3[!(housing2_raw$LotConfig == "CulDSac" | housing2_raw$LotConfig == "FR3")] <- 0

table(housing2_raw$LotConfig_culdsac_fr3)
housing2_raw$LotConfig <- NULL

##8 LandSlope
summary(housing2_raw$LandSlope)

### We don't see much difference with this. So we will drop this feature.
housing2_raw$LandSlope <- NULL

##9 Neighborhood
summary(housing2_raw$Neighborhood)

housing2_raw$Neighborhood_highprice <- ifelse(housing2_raw$Neighborhood %in% 
                                                (which(with(housing2_raw, 
                                                            tapply(SalePrice, Neighborhood, median)) > 200000) %>% names()), 
                                              1, 0)

table(housing2_raw$Neighborhood_highprice)

housing2_raw$Neighborhood <- NULL

##10 Condition1
summary(housing2_raw$Condition1)

housing2_raw$Condition1_good[housing2_raw$Condition1 == "PosN" | housing2_raw$Condition1 == "PosA"] <- 1
housing2_raw$Condition1_good[!(housing2_raw$Condition1 == "PosN" | housing2_raw$Condition1 == "PosA")] <- 0

table(housing2_raw$Condition1_good)
housing2_raw$Condition1 <- NULL

##11 Condition2
summary(housing2_raw$Condition2)

housing2_raw$Condition2_good[housing2_raw$Condition2 == "PosN" | housing2_raw$Condition2 == "PosA"] <- 1
housing2_raw$Condition2_good[!(housing2_raw$Condition2 == "PosN" | housing2_raw$Condition2 == "PosA")] <- 0

table(housing2_raw$Condition2_good)
housing2_raw$Condition2 <- NULL

##12 BldgType
summary(housing2_raw$BldgType)

housing2_raw$BldgType_singlefam_endtwnhse[housing2_raw$BldgType == "1Fam" | housing2_raw$BldgType == "TwnhsE"] <- 1
housing2_raw$BldgType_singlefam_endtwnhse[!(housing2_raw$BldgType == "1Fam" | housing2_raw$BldgType == "TwnhsE")] <- 0

table(housing2_raw$BldgType_singlefam_endtwnhse)
housing2_raw$BldgType <- NULL

##13 HouseStyle
summary(housing2_raw$HouseStyle)

housing2_raw$HouseStyle_level[housing2_raw$HouseStyle == "2Story" | housing2_raw$HouseStyle == "2.5Fin"] <- 1
housing2_raw$HouseStyle_level[!(housing2_raw$HouseStyle == "2Story" | housing2_raw$HouseStyle == "2.5Fin")] <- 0

table(housing2_raw$HouseStyle_level)
housing2_raw$HouseStyle <- NULL

##14 RoofStyle
summary(housing2_raw$RoofStyle)

housing2_raw$RoofStyle_level[housing2_raw$RoofStyle == "Hip" | housing2_raw$RoofStyle == "Flat" | housing2_raw$RoofStyle == "Shed"] <- 1
housing2_raw$RoofStyle_level[!(housing2_raw$RoofStyle == "Hip" | housing2_raw$RoofStyle == "Flat" | housing2_raw$RoofStyle == "Shed")] <- 0

table(housing2_raw$RoofStyle_level)
housing2_raw$RoofStyle <- NULL

##15 RoofMatl
summary(housing2_raw$RoofMatl)

housing2_raw$RoofMatl_level[housing2_raw$RoofMatl == "WdShake" | housing2_raw$RoofMatl == "WdShngl" | housing2_raw$RoofMatl == "Membran"] <- 1
housing2_raw$RoofMatl_level[!(housing2_raw$RoofMatl == "WdShake" | housing2_raw$RoofMatl == "WdShngl" | housing2_raw$RoofMatl == "Membran")] <- 0

table(housing2_raw$RoofMatl_level)
housing2_raw$RoofMatl <- NULL

##16 and 17Exterior1st and Exterior2nd

summary(housing2_raw$Exterior1st)
summary(housing2_raw$Exterior2nd)

## Not using these as there is not much difference in the SalePrice
housing2_raw$Exterior1st <- NULL
housing2_raw$Exterior2nd <- NULL

#18 MasVnrType
summary(housing2_raw$MasVnrType)

housing2_raw$MasVnrType_Stone[housing2_raw$MasVnrType == "Stone" & !is.na(housing2_raw$MasVnrType)] <- 1
housing2_raw$MasVnrType_Stone[!(housing2_raw$MasVnrType == "Stone" & !is.na(housing2_raw$MasVnrType))] <- 0

table(housing2_raw$MasVnrType_Stone)
housing2_raw$MasVnrType <- NULL

#19 ExterQual
summary(housing2_raw$ExterQual)

housing2_raw$ExterQual_level[housing2_raw$ExterQual == "Fa"] <- 1
housing2_raw$ExterQual_level[housing2_raw$ExterQual == "TA"] <- 2
housing2_raw$ExterQual_level[housing2_raw$ExterQual == "Gd"] <- 3
housing2_raw$ExterQual_level[housing2_raw$ExterQual == "Ex"] <- 4

table(housing2_raw$ExterQual_level)
housing2_raw$ExterQual <- NULL

#20 ExterCond
summary(housing2_raw$ExterCond)

housing2_raw$ExterCond_level[housing2_raw$ExterCond == "Po"] <- 1
housing2_raw$ExterCond_level[housing2_raw$ExterCond == "Fa"] <- 2
housing2_raw$ExterCond_level[housing2_raw$ExterCond == "TA"] <- 3
housing2_raw$ExterCond_level[housing2_raw$ExterCond == "Gd"] <- 4
housing2_raw$ExterCond_level[housing2_raw$ExterCond == "Ex"] <- 5

table(housing2_raw$ExterCond_level)
housing2_raw$ExterCond <- NULL

#21 Foundation
summary(housing2_raw$Foundation)

housing2_raw$Foundation_concrete[housing2_raw$Foundation == "PConc"] <- 1
housing2_raw$Foundation_concrete[!(housing2_raw$Foundation == "PConc")] <- 0

table(housing2_raw$Foundation_concrete)
housing2_raw$Foundation <- NULL

#22 BsmtQual
summary(housing2_raw$BsmtQual)


housing2_raw$BsmtQual_level[housing2_raw$BsmtQual == "Fa" & !is.na(housing2_raw$BsmtQual)] <- 2
housing2_raw$BsmtQual_level[housing2_raw$BsmtQual == "TA" & !is.na(housing2_raw$BsmtQual)] <- 3
housing2_raw$BsmtQual_level[housing2_raw$BsmtQual == "Gd" & !is.na(housing2_raw$BsmtQual)] <- 4
housing2_raw$BsmtQual_level[housing2_raw$BsmtQual == "Ex" & !is.na(housing2_raw$BsmtQual)] <- 5
housing2_raw$BsmtQual_level[is.na(housing2_raw$BsmtQual)] <- 1
housing2_raw$BsmtQual <- NULL

#23 BsmtCond
summary(housing2_raw$BsmtCond)

housing2_raw$BsmtCond_level[housing2_raw$BsmtCond == "Po" | is.na(housing2_raw$BsmtCond)] <- 1
housing2_raw$BsmtCond_level[housing2_raw$BsmtCond == "Fa" & !is.na(housing2_raw$BsmtCond)] <- 2
housing2_raw$BsmtCond_level[housing2_raw$BsmtCond == "TA" & !is.na(housing2_raw$BsmtCond)] <- 3
housing2_raw$BsmtCond_level[housing2_raw$BsmtCond == "Gd" & !is.na(housing2_raw$BsmtCond)] <- 4
housing2_raw$BsmtCond_level[housing2_raw$BsmtCond == "Ex" & !is.na(housing2_raw$BsmtCond)] <- 5

housing2_raw$BsmtCond <- NULL

#24 BsmtExposure
summary(housing2_raw$BsmtExposure)

housing2_raw$BsmtExposure_level[housing2_raw$BsmtExposure == "No" | is.na(housing2_raw$BsmtExposure)] <- 1
housing2_raw$BsmtExposure_level[housing2_raw$BsmtExposure == "Mn" & !is.na(housing2_raw$BsmtExposure)] <- 2
housing2_raw$BsmtExposure_level[housing2_raw$BsmtExposure == "Av" & !is.na(housing2_raw$BsmtExposure)] <- 3
housing2_raw$BsmtExposure_level[housing2_raw$BsmtExposure == "Gd" & !is.na(housing2_raw$BsmtExposure)] <- 4

table(housing2_raw$BsmtExposure_level)
housing2_raw$BsmtExposure <- NULL

## 25 and 26 - BsmtFinType1 and BsmtFinType2
### Not using these features as we have already considered the most important Basement
### features
housing2_raw$BsmtFinType1 <- NULL
housing2_raw$BsmtFinType2 <- NULL

## 27 - Heating
summary(housing2_raw$Heating)

housing2_raw$Heating_type[housing2_raw$Heating == "GasW" | housing2_raw$Heating == "GasA"] <- 1
housing2_raw$Heating_type[!(housing2_raw$Heating == "GasW" | housing2_raw$Heating == "GasA")] <- 0

table(housing2_raw$Heating_type)
housing2_raw$Heating <- NULL

## 28 - hEATINGqc
summary(housing2_raw$HeatingQC)

housing2_raw$HeatingQC_level[housing2_raw$HeatingQC == "Po"] <- 1
housing2_raw$HeatingQC_level[housing2_raw$HeatingQC == "Fa"] <- 2
housing2_raw$HeatingQC_level[housing2_raw$HeatingQC == "TA"] <- 3
housing2_raw$HeatingQC_level[housing2_raw$HeatingQC == "Gd"] <- 4
housing2_raw$HeatingQC_level[housing2_raw$HeatingQC == "Ex"] <- 5

table(housing2_raw$HeatingQC_level)
housing2_raw$HeatingQC <- NULL

##29 - CentralAir
summary(housing2_raw$CentralAir)

housing2_raw$CentralAir_flag[housing2_raw$CentralAir == "Y"] <- 1
housing2_raw$CentralAir_flag[housing2_raw$CentralAir == "N"] <- 0

table(housing2_raw$CentralAir_flag)
housing2_raw$CentralAir <- NULL

##30 - Electrical
summary(housing2_raw$Electrical)

housing2_raw$Electrical_SBrkr[housing2_raw$Electrical == "SBrkr" & !is.na(housing2_raw$Electrical)] <- 1
housing2_raw$Electrical_SBrkr[!(housing2_raw$Electrical == "SBrkr" & !is.na(housing2_raw$Electrical))] <- 0

table(housing2_raw$Electrical_SBrkr)
housing2_raw$Electrical <- NULL

##31 - KitchenQual
summary(housing2_raw$KitchenQual)

housing2_raw$KitchenQual_level[housing2_raw$KitchenQual == "Fa"] <- 1
housing2_raw$KitchenQual_level[housing2_raw$KitchenQual == "TA"] <- 2
housing2_raw$KitchenQual_level[housing2_raw$KitchenQual == "Gd"] <- 3
housing2_raw$KitchenQual_level[housing2_raw$KitchenQual == "Ex"] <- 4

table(housing2_raw$KitchenQual_level)
housing2_raw$KitchenQual <- NULL

##32 - fUNCTIONAL
## As seen in the box plots, we don't see much differences in the various category levels
## So we won't be using this feature

housing2_raw$Functional <- NULL

##33 - FireplaceQu
summary(housing2_raw$FireplaceQu)

housing2_raw$FireplaceQu_level[housing2_raw$FireplaceQu == "Po" | is.na(housing2_raw$FireplaceQu)] <- 1
housing2_raw$FireplaceQu_level[housing2_raw$FireplaceQu == "Fa" & !is.na(housing2_raw$FireplaceQu)] <- 2
housing2_raw$FireplaceQu_level[housing2_raw$FireplaceQu == "TA" & !is.na(housing2_raw$FireplaceQu)] <- 3
housing2_raw$FireplaceQu_level[housing2_raw$FireplaceQu == "Gd" & !is.na(housing2_raw$FireplaceQu)] <- 4
housing2_raw$FireplaceQu_level[housing2_raw$FireplaceQu == "Ex" & !is.na(housing2_raw$FireplaceQu)] <- 5

table(housing2_raw$FireplaceQu_level)
housing2_raw$FireplaceQu <- NULL

##34 - GarageType
summary(housing2_raw$GarageType)

housing2_raw$GarageType_within[(housing2_raw$GarageType == "Attchd" | housing2_raw$GarageType == "BuiltIn") & !is.na(housing2_raw$GarageType)] <- 1
housing2_raw$GarageType_within[!((housing2_raw$GarageType == "Attchd" | housing2_raw$GarageType == "BuiltIn") & !is.na(housing2_raw$GarageType))] <- 0

table(housing2_raw$GarageType_within)
housing2_raw$GarageType <- NULL

##35 - gARAGEfINISH
summary(housing2_raw$GarageFinish)

housing2_raw$GarageFinish_status[(housing2_raw$GarageFinish == "Fin" | housing2_raw$GarageFinish == "RFn") & !is.na(housing2_raw$GarageFinish)] <- 1
housing2_raw$GarageFinish_status[!((housing2_raw$GarageFinish == "Fin" | housing2_raw$GarageFinish == "RFn") & !is.na(housing2_raw$GarageFinish))] <- 0

table(housing2_raw$GarageFinish_status)
housing2_raw$GarageFinish <- NULL

##36 - GarageQual
summary(housing2_raw$GarageQual)

housing2_raw$GarageQual_level[housing2_raw$GarageQual == "Po" | is.na(housing2_raw$GarageQual)] <- 1
housing2_raw$GarageQual_level[housing2_raw$GarageQual == "Fa" & !is.na(housing2_raw$GarageQual)] <- 2
housing2_raw$GarageQual_level[housing2_raw$GarageQual == "TA" & !is.na(housing2_raw$GarageQual)] <- 3
housing2_raw$GarageQual_level[housing2_raw$GarageQual == "Gd" & !is.na(housing2_raw$GarageQual)] <- 4
housing2_raw$GarageQual_level[housing2_raw$GarageQual == "Ex" & !is.na(housing2_raw$GarageQual)] <- 5

table(housing2_raw$GarageQual_level)
housing2_raw$GarageQual <- NULL

##37 - GarageCond
summary(housing2_raw$GarageCond)

housing2_raw$GarageCond_status[housing2_raw$GarageCond == "Po" | is.na(housing2_raw$GarageCond)] <- 1
housing2_raw$GarageCond_status[housing2_raw$GarageCond == "Fa" & !is.na(housing2_raw$GarageCond)] <- 2
housing2_raw$GarageCond_status[housing2_raw$GarageCond == "TA" & !is.na(housing2_raw$GarageCond)] <- 3
housing2_raw$GarageCond_status[housing2_raw$GarageCond == "Gd" & !is.na(housing2_raw$GarageCond)] <- 4
housing2_raw$GarageCond_status[housing2_raw$GarageCond == "Ex" & !is.na(housing2_raw$GarageCond)] <- 5

table(housing2_raw$GarageCond_status)
housing2_raw$GarageCond <- NULL

##38 - PavedDrive
summary(housing2_raw$PavedDrive)

housing2_raw$PavedDrive_flag[housing2_raw$PavedDrive == "Y" | housing2_raw$PavedDrive == "P"] <- 1
housing2_raw$PavedDrive_flag[housing2_raw$PavedDrive == "N"] <- 0

table(housing2_raw$PavedDrive_flag)
housing2_raw$PavedDrive <- NULL

##39 - PoolQC
summary(housing2_raw$PoolQC)

housing2_raw$PoolQC_level[housing2_raw$PoolQC == "Fa" | is.na(housing2_raw$PoolQC)] <- 1
housing2_raw$PoolQC_level[housing2_raw$PoolQC == "Gd" & !is.na(housing2_raw$PoolQC)] <- 2
housing2_raw$PoolQC_level[housing2_raw$PoolQC == "Ex" & !is.na(housing2_raw$PoolQC)] <- 3

table(housing2_raw$PoolQC_level)
housing2_raw$PoolQC <- NULL

##40 - Fence
summary(housing2_raw$Fence)

housing2_raw$Fence_good[housing2_raw$Fence == "GdPrv" & !is.na(housing2_raw$Fence)] <- 1
housing2_raw$Fence_good[!(housing2_raw$Fence == "GdPrv" & !is.na(housing2_raw$Fence))] <- 0

table(housing2_raw$Fence_good)
housing2_raw$Fence <- NULL

##41 - MiscFeature
### nOT MUCH difference in the saleprice in the levels, hence we won't use it
housing2_raw$MiscFeature <- NULL

##42 - MoSold
### nOT MUCH difference in the saleprice in the levels, hence we won't use it
housing2_raw$MoSold <- NULL

##43 - sALEtYPE
summary(housing2_raw$SaleType)

housing2_raw$SaleType_value[housing2_raw$SaleType == "New" | housing2_raw$SaleType == "Con"] <- 3
housing2_raw$SaleType_value[housing2_raw$SaleType == "ConLw" | housing2_raw$SaleType == "CWD" | housing2_raw$SaleType == "WD"] <- 2
housing2_raw$SaleType_value[housing2_raw$SaleType == "COD" | housing2_raw$SaleType == "ConLD" | 
                        housing2_raw$SaleType == "ConLI"  | housing2_raw$SaleType == "Oth"] <- 1

table(housing2_raw$SaleType_value)

housing2_raw$SaleType <- NULL

## 44 - SaleCondition
### Dropping this feature and not using it
housing2_raw$SaleCondition <- NULL

########## All features are now ready
dim(housing2_raw)
summary(housing2_raw)


```


Checking Missing data again - to ensure all good, and there is no missing data 
```{r}
### Blank data
train_missing_df2 <- data.frame(apply(housing2_raw, 2, function(x) length(which(x == ''))))
colnames(train_missing_df2) <- "counts"
train_missing_df2$column_name <- rownames(train_missing_df2)
train_missing_df2 <- data.frame(train_missing_df2[train_missing_df2$counts > 0,])
print(train_missing_df2)

### NA data
train_na_df2 <- data.frame(apply(housing2_raw, 2, function(x) length(which(is.na(x)))))

colnames(train_na_df2) <- "counts"
train_na_df2$column_name <- rownames(train_na_df2)
train_na_df2 <- data.frame(train_na_df2[train_na_df2$counts > 0,])
print(train_na_df2)

```

Now there is no missing data or NA's.

#### Preparing for the model now.
Taking the SalePrice - the target variable to the front of the dataframe by reordering the features in the dataframe.
```{r}
housing2_raw <- housing2_raw[,c(36,1:35,37:70)]

colnames(housing2_raw)
```

##### Test-Train split
Splitting the data into test and train datasets
```{r}
n <- nrow(housing2_raw)
set.seed(123)
housing2_random_index <- housing2_raw[sample(n), ]

housing2.train.df <- housing2_raw[1:as.integer(0.8*n),]

housing2.test.df <- housing2_raw[as.integer(0.8*n +1):n, ]

nrow(housing2.train.df)
nrow(housing2.test.df)


```

#### Models
Building first model - taking all the features.
```{r}
#### Linear Model 1
g1 <- lm(SalePrice ~ ., 
         data = housing2.train.df)

summary(g1) 
```

As we see above, many of the variables are not statistically significant. So we are only taking the significant features.
```{r}
g2 <- lm(SalePrice ~ MSSubClass + LotFrontage + LotArea + OverallQual + OverallCond +
           YearBuilt + MasVnrArea + BsmtFinSF1 + BsmtFinSF2 + BsmtUnfSF + X1stFlrSF +
           X2ndFlrSF + LowQualFinSF + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd +
           GarageArea + WoodDeckSF + ScreenPorch + Paved_alley + LotConfig_culdsac_fr3 +
           Neighborhood_highprice + Condition1_good + Condition2_good + RoofMatl_level +
           ExterQual_level + BsmtQual_level + BsmtCond_level + BsmtExposure_level +
           KitchenQual_level + PoolQC_level + SaleType_value + GarageCond_status, 
         data = housing2.train.df)

summary(g2) 

```
Model diagnostic plots
```{r}
#resid_panel(g2, plots='default', smoother = TRUE)
plot(g2)
```


Removing statistically insignificant features further.
```{r}
g3 <- lm(SalePrice ~ MSSubClass + LotFrontage + LotArea + OverallQual + OverallCond +
           YearBuilt + MasVnrArea + BsmtFinSF1 + BsmtFinSF2 + BsmtUnfSF + X1stFlrSF +
           X2ndFlrSF + LowQualFinSF + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd +
           GarageArea + WoodDeckSF + ScreenPorch +
           Neighborhood_highprice + Condition1_good + Condition2_good + RoofMatl_level +
           ExterQual_level + BsmtQual_level + BsmtCond_level + BsmtExposure_level +
           KitchenQual_level + PoolQC_level + SaleType_value + GarageCond_status, 
         data = housing2.train.df)

summary(g3) 

```

Model diagnostic plots
```{r}
#resid_panel(g3, plots='default', smoother = TRUE)
plot(g3)
```

As residuals are a curve shape, we will try taking the log of the SalePrice, which is the target variable.
```{r}
### Taking log
g4 <- lm(log(SalePrice) ~ MSSubClass + LotFrontage + LotArea + OverallQual +
           OverallCond +
           YearBuilt + MasVnrArea + BsmtFinSF1 + BsmtFinSF2 + BsmtUnfSF + X1stFlrSF +
           X2ndFlrSF + LowQualFinSF + BedroomAbvGr + KitchenAbvGr + TotRmsAbvGrd +
           GarageArea + WoodDeckSF + ScreenPorch +
           Neighborhood_highprice + Condition1_good + Condition2_good + RoofMatl_level
         + ExterQual_level + BsmtQual_level + BsmtCond_level + BsmtExposure_level +
           KitchenQual_level + PoolQC_level + SaleType_value + GarageCond_status, 
         data = housing2.train.df)

summary(g4) 

```

Model diagnostic plots
```{r}
plot(g4)
```

Removing further the features which are statistically insignificant
```{r}
## Removing further 
g6 <- lm(log(SalePrice) ~ MSSubClass + LotFrontage + LotArea + OverallQual + OverallCond +
           YearBuilt + BsmtFinSF1 + BsmtFinSF2 + BsmtUnfSF + X1stFlrSF +
           X2ndFlrSF + LowQualFinSF + KitchenAbvGr + TotRmsAbvGrd +
           GarageArea + WoodDeckSF + ScreenPorch +
           Neighborhood_highprice + Condition2_good +
           BsmtQual_level + BsmtExposure_level +
           KitchenQual_level + SaleType_value + GarageCond_status, 
         data = housing2.train.df)

summary(g6) 

```

Model diagnostic plots
```{r}
plot(g6)
```


This is the best one as it has the biggest R-squared, and the plots also suggest that the assumptions of the linear model are all true.
###### Prediction of test dataset
```{r}
Predicted_test_raw2 <- predict(g6, newdata = housing2.test.df)

Predicted_test_final2 <- exp(Predicted_test_raw2)

Test_predicted_values2 <- cbind(data.frame(housing2.test.df$SalePrice), 
                               data.frame(Predicted_test_final2))

class(Test_predicted_values2)
names(Test_predicted_values2) <- c("Actual", "Predicted")

head(Test_predicted_values2, 20)

```

Plots for the most relevant features - one by one:

```{r}
summary(housing2_raw$OverallQual)
ggplot(data = housing2_raw, aes(x = as.factor(OverallQual), y = SalePrice)) +
  geom_boxplot()

summary(housing2_raw$OverallCond)
ggplot(data = housing2_raw, aes(x = as.factor(OverallCond), y = SalePrice)) +
  geom_boxplot()

summary(housing2_raw$YearBuilt)
ggplot(data = housing2_raw, aes(x = YearBuilt, y = SalePrice)) +
  geom_point() + geom_smooth(method=lm , color="blue", se=FALSE)

summary(housing2_raw$BsmtFinSF1)
ggplot(data = housing2_raw, aes(x = BsmtFinSF1, y = SalePrice)) +
  geom_point() + geom_smooth(method=lm , color="blue", se=FALSE)

summary(housing2_raw$X1stFlrSF)
ggplot(data = housing2_raw, aes(x = X1stFlrSF, y = SalePrice)) +
  geom_point() + geom_smooth(method=lm , color="blue", se=FALSE)

summary(housing2_raw$X2ndFlrSF)
ggplot(data = housing2_raw, aes(x = X2ndFlrSF, y = SalePrice)) +
  geom_point() + geom_smooth(method=lm , color="blue", se=FALSE)


summary(housing2_raw$Neighborhood_highprice)
ggplot(data = housing2_raw, aes(x = as.factor(Neighborhood_highprice), 
                                y = SalePrice)) +
  geom_boxplot()

summary(housing2_raw$Condition2_good)
ggplot(data = housing2_raw, aes(x = as.factor(Condition2_good), 
                                y = SalePrice)) +
  geom_boxplot()

```


#### Conclusion
1. There is a very clear Log linear relationship between the target feature - Sale Price and the following independent features. 

MSSubClass

LotFrontage

LotArea

OverallQual

OverallCond

YearBuilt

BsmtFinSF1

BsmtFinSF2

BsmtUnfSF

X1stFlrSF

X2ndFlrSF

LowQualFinSF

KitchenAbvGr

TotRmsAbvGrd

GarageArea

WoodDeckSF

ScreenPorch

Neighborhood_highprice

Condition2_good

BsmtQual_level

BsmtExposure_level

KitchenQual_level

SaleType_value

GarageCond_status


2. The most relevant features are:

OverallQual

OverallCond

YearBuilt

BsmtFinSF1

X1stFlrSF

X2ndFlrSF

Condition2_good
