---
title: "Data621_HW_4_Car_Insurance"
author: "Deepak Mongia"
date: "4/13/2020"
output: html_document
---

In this homework assignment, we will explore, analyze and model a data set containing approximately 8000 records representing a customer at an auto insurance company. Each record has two response variables. The first response variable, TARGET_FLAG, is a 1 or a 0. A “1” means that the person was in a car crash. A zero means that the person was not in a car crash. The second response variable is TARGET_AMT. This value is zero if the person did not crash their car. But if they did crash their car this number will be a value greater than zero.

Our objective is to build multiple linear regression and binary logistic regression models on the training data to predict the probability that a person will crash their car and also the amount of money it will cost if the person does crash their car.

Loading the libraries:
```{r}
library(dplyr)
library(ggplot2)
library(stringr)
library(mice)
library(VIM)
library(tidyr)
library(pROC)
```

Loading the data, and exploring the data:
```{r}
insurance_raw <- read.csv("https://raw.githubusercontent.com/deepakmongia/Data621/master/HW-4/Data/insurance_training_data.csv",
                          header = TRUE, row.names = 1)

print(head(insurance_raw))

print(dim(insurance_raw))

print(str(insurance_raw))
print(summary(insurance_raw))

insurance_raw$TARGET_FLAG <- as.factor(insurance_raw$TARGET_FLAG)

ggplot(insurance_raw, aes(TARGET_FLAG)) + geom_bar(aes(fill=TARGET_FLAG))

print(table(insurance_raw$TARGET_FLAG))

```

As we see above the people which have claimed (target flag = 1) are 2153 in number and who have not claimed (target flag = 0) are 6008. So it is an imbalanced dataset.

### Missing data checks:
```{r}
### Blank data
train_missing_df <- data.frame(apply(insurance_raw, 2, function(x) length(which(x == ''))))
print(train_missing_df)

### NA data
train_na_df1 <- data.frame(apply(insurance_raw, 2, function(x) length(which(is.na(x)))))
print(train_na_df1)

```


### Analyzing the independent features

##### KIDSDRIV
```{r}
print(unique(insurance_raw$KIDSDRIV))

ggplot(insurance_raw, aes(KIDSDRIV)) + geom_bar(aes(fill=KIDSDRIV))

insurance_raw[insurance_raw$KIDSDRIV != 0,] %>% ggplot(aes(KIDSDRIV)) + 
  geom_bar(aes(fill=KIDSDRIV))

Adults_only_policies <- insurance_raw[insurance_raw$KIDSDRIV == 0,]
with_kids_policies <- insurance_raw[insurance_raw$KIDSDRIV != 0,]

table(Adults_only_policies$TARGET_FLAG) / nrow(Adults_only_policies)
table(with_kids_policies$TARGET_FLAG) / nrow(with_kids_policies)

```

##### AGE
```{r}
### Age
ggplot(insurance_raw, aes(AGE)) + geom_histogram(binwidth = 5)

ggplot(insurance_raw, aes(AGE)) + geom_histogram(binwidth = 5) +
  facet_grid(~TARGET_FLAG)

ggplot(insurance_raw, aes(x = AGE, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

```

##### HOMEKIDS
```{r}
### homekids
unique(insurance_raw$HOMEKIDS)

kids_0_insurance <- insurance_raw[insurance_raw$HOMEKIDS == 0,]
kids_non0_insurance <- insurance_raw[insurance_raw$HOMEKIDS != 0,]


kids_0_insurance$TARGET_FLAG <- as.integer(as.character(kids_0_insurance$TARGET_FLAG))
kids_non0_insurance$TARGET_FLAG <- as.integer(as.character(kids_non0_insurance$TARGET_FLAG))

ggplot(kids_0_insurance, aes(TARGET_FLAG)) + geom_histogram()
ggplot(kids_non0_insurance, aes(TARGET_FLAG)) + geom_histogram()

table(kids_0_insurance$TARGET_FLAG) / nrow(kids_0_insurance)

table(kids_non0_insurance$TARGET_FLAG) / nrow(kids_non0_insurance)

ggplot(insurance_raw, aes(x = HOMEKIDS, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

```

##### years on job
```{r}
### Years on job
ggplot(data = insurance_raw, aes(x = YOJ, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

ggplot(insurance_raw, aes(x = YOJ, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")


```

##### INCOME
```{r}
### INCOME
any(is.na(insurance_raw$INCOME))

any(insurance_raw$INCOME == '')

insurance_raw$INCOME[insurance_raw$INCOME == ''] <- NA

sum(is.na(insurance_raw$INCOME))

insurance_raw$INCOME <- str_remove_all(insurance_raw$INCOME, "[$,]") %>% as.integer()

ggplot(insurance_raw, aes(INCOME)) + geom_histogram() +
  facet_grid(~TARGET_FLAG)

ggplot(data = insurance_raw, aes(x = INCOME, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

```

##### PARENT1
```{r}
### PARENT1
xtabs(~insurance_raw$PARENT1 + insurance_raw$TARGET_FLAG)

```

##### HOME_VAL
```{r}
sum(is.na(insurance_raw$HOME_VAL))

insurance_raw$HOME_VAL[insurance_raw$HOME_VAL == ''] <- NA

sum(is.na(insurance_raw$HOME_VAL))

insurance_raw$HOME_VAL <- str_remove_all(insurance_raw$HOME_VAL, "[$,]") %>% as.integer()

ggplot(insurance_raw, aes(HOME_VAL)) + geom_histogram() +
  facet_grid(~TARGET_FLAG)

ggplot(data = insurance_raw, aes(x = HOME_VAL, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()


```

##### Marital status
```{r}
unique(insurance_raw$MSTATUS)

ggplot(data = insurance_raw, aes(x = MSTATUS)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

insurance_raw %>% 
  ggplot(aes(MSTATUS)) +
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())


```

##### sex
```{r}

ggplot(data = insurance_raw, aes(x = SEX)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())


```

##### education
```{r}
unique(insurance_raw$EDUCATION)

ggplot(data = insurance_raw, aes(x = EDUCATION)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

```

##### job
```{r}
unique(insurance_raw$JOB)

insurance_raw$JOB[insurance_raw$JOB == ''] <- NA

ggplot(data = insurance_raw, aes(x = JOB)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

```

##### TRAVTIME
```{r}

ggplot(insurance_raw, aes(x = TRAVTIME, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

ggplot(data = insurance_raw, aes(x = TRAVTIME, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

```

##### CAR_USE
```{r}
unique(insurance_raw$CAR_USE)

ggplot(data = insurance_raw, aes(x = CAR_USE)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())


```

##### BLUEBOOK
```{r}
sum(insurance_raw$BLUEBOOK == '')

insurance_raw$BLUEBOOK <- str_remove_all(insurance_raw$BLUEBOOK, "[$,]") %>% as.integer()

#ggplot(insurance_raw, aes(BLUEBOOK)) + geom_histogram() +
#  facet_grid(~TARGET_FLAG)

#ggplot(insurance_raw, aes(x = BLUEBOOK, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
#  geom_histogram()

ggplot(insurance_raw, aes(x = BLUEBOOK, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

ggplot(data = insurance_raw, aes(x = BLUEBOOK, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()
```

##### Time in force - TIF
```{r}

ggplot(insurance_raw, aes(x = TIF, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

ggplot(data = insurance_raw, aes(x = TIF, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

```

##### CAR_TYPE
```{r}

unique(insurance_raw$CAR_TYPE)

ggplot(data = insurance_raw, aes(x = CAR_TYPE)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

```

##### RED_CAR
```{r}

unique(insurance_raw$RED_CAR)

ggplot(data = insurance_raw, aes(x = RED_CAR)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

RED_CAR_xtabs <- xtabs(~insurance_raw$RED_CAR + insurance_raw$TARGET_FLAG)

print(RED_CAR_xtabs)

RED_CAR_xtabs[1,] <- RED_CAR_xtabs[1,] / sum(RED_CAR_xtabs[1,])
RED_CAR_xtabs[2,] <- RED_CAR_xtabs[2,] / sum(RED_CAR_xtabs[2,])
print(RED_CAR_xtabs)

```

#####OLDCLAIM
```{r}
sum(insurance_raw$OLDCLAIM == '')

insurance_raw$OLDCLAIM <- str_remove_all(insurance_raw$OLDCLAIM, "[$,]") %>% as.integer()

ggplot(insurance_raw, aes(x = OLDCLAIM, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

ggplot(data = insurance_raw, aes(x = OLDCLAIM, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

ggplot(data = insurance_raw, aes(x = OLDCLAIM, y = TARGET_AMT)) + geom_point()


```

```{r}
### CLM_FREQ
unique(insurance_raw$CLM_FREQ)

ggplot(data = insurance_raw, aes(x = CLM_FREQ)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

CLM_FREQ_xtabs <- xtabs(~insurance_raw$CLM_FREQ + insurance_raw$TARGET_FLAG)

print(CLM_FREQ_xtabs)

CLM_FREQ_xtabs[1,] <- CLM_FREQ_xtabs[1,] / sum(CLM_FREQ_xtabs[1,])
CLM_FREQ_xtabs[2,] <- CLM_FREQ_xtabs[2,] / sum(CLM_FREQ_xtabs[2,])
CLM_FREQ_xtabs[3,] <- CLM_FREQ_xtabs[3,] / sum(CLM_FREQ_xtabs[3,])
CLM_FREQ_xtabs[4,] <- CLM_FREQ_xtabs[4,] / sum(CLM_FREQ_xtabs[4,])
CLM_FREQ_xtabs[5,] <- CLM_FREQ_xtabs[5,] / sum(CLM_FREQ_xtabs[5,])
CLM_FREQ_xtabs[6,] <- CLM_FREQ_xtabs[6,] / sum(CLM_FREQ_xtabs[6,])

print(CLM_FREQ_xtabs)

```

```{r}
### REVOKED
unique(insurance_raw$REVOKED)

ggplot(data = insurance_raw, aes(x = REVOKED)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

REVOKED_xtabs <- xtabs(~insurance_raw$REVOKED + insurance_raw$TARGET_FLAG)

print(REVOKED_xtabs)

REVOKED_xtabs[1,] <- REVOKED_xtabs[1,] / sum(REVOKED_xtabs[1,])
REVOKED_xtabs[2,] <- REVOKED_xtabs[2,] / sum(REVOKED_xtabs[2,])
print(REVOKED_xtabs)

```

##### MVR_PTS
```{r}

print(summary(insurance_raw$MVR_PTS))
print(unique(insurance_raw$MVR_PTS))

ggplot(insurance_raw, aes(x = MVR_PTS, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

ggplot(data = insurance_raw, aes(x = MVR_PTS, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

```

##### CAR_AGE
```{r}

summary(insurance_raw$CAR_AGE)

ggplot(insurance_raw, aes(x = CAR_AGE, group = TARGET_FLAG, fill = TARGET_FLAG)) + 
  geom_histogram(position = "dodge")

ggplot(data = insurance_raw, aes(x = CAR_AGE, y = TARGET_FLAG)) + geom_point() +
  geom_jitter()

```

```{r}
### urbanicity

print(summary(insurance_raw$URBANICITY))
print(unique(insurance_raw$URBANICITY))


ggplot(data = insurance_raw, aes(x = URBANICITY)) + 
  geom_bar(aes(fill = TARGET_FLAG), position = position_dodge())

URBANICITY_xtabs <- xtabs(~insurance_raw$URBANICITY + insurance_raw$TARGET_FLAG)

print(URBANICITY_xtabs)

URBANICITY_xtabs[1,] <- URBANICITY_xtabs[1,] / sum(URBANICITY_xtabs[1,])
URBANICITY_xtabs[2,] <- URBANICITY_xtabs[2,] / sum(URBANICITY_xtabs[2,])
print(URBANICITY_xtabs)

```

#### Imputing the missing data
We will impute only the numerical missing data. We also have 1 feature which is categorical and missing some data. We will not impute the missing categorical data - feature JOB, and instead we will drop the records having NA in this feature.
Rest all missing data is the numerical field data, and we are using the mice package below to impute it.
```{r}

#sum(!complete.cases(insurance_raw))

summary(insurance_raw$YOJ)

summary(insurance_raw$AGE)

summary(insurance_raw$INCOME)

summary(insurance_raw$HOME_VAL)

summary(insurance_raw$JOB)

summary(insurance_raw$CAR_AGE)


md.pattern(insurance_raw)

insurance_raw2 <- insurance_raw %>% drop_na(JOB)

aggr_plot <- aggr(insurance_raw2, col=c('navyblue','red'), 
                  numbers=TRUE, sortVars=TRUE, 
                  labels=names(insurance_raw), cex.axis=.7, 
                  gap=3, ylab=c("Histogram of missing data","Pattern"))



insurance_raw_imputed <- mice(data = insurance_raw2, m = 1,
                              method = "pmm", maxit = 5, seed = 500)

insurance_raw_imputed_df <- mice::complete(insurance_raw_imputed, 1)

head(insurance_raw_imputed_df)

summary(insurance_raw$YOJ)
summary(insurance_raw_imputed_df$YOJ)

summary(insurance_raw$AGE)
summary(insurance_raw_imputed_df$AGE)

summary(insurance_raw$INCOME)
summary(insurance_raw_imputed_df$INCOME)

summary(insurance_raw$HOME_VAL)
summary(insurance_raw_imputed_df$HOME_VAL)

summary(insurance_raw$JOB)
summary(insurance_raw_imputed_df$JOB)

summary(insurance_raw$CAR_AGE)
summary(insurance_raw_imputed_df$CAR_AGE)

```

Now all the data is present. So we can start building our models.

### Model Building
#### Train-test Split
We are splitting our data into training and test dataset - first for our classification  problem.
```{r}
#### Splitting the dataset into test and train dataset
n <- nrow(insurance_raw_imputed_df)
set.seed(123)
insurance_random_index <- insurance_raw_imputed_df[sample(n), ]

insurance.train.df <- insurance_random_index[1:as.integer(0.7*n),]

insurance.test.df <- insurance_random_index[as.integer(0.7*n +1):n, ]

table(insurance.test.df$TARGET_FLAG) / nrow(insurance.test.df)
table(insurance.train.df$TARGET_FLAG) / nrow(insurance.train.df)

```

#### Building our first classification model

Taking all features
```{r}
### Building the first Logistic Regression model using all the variables:

logitModel1 <- glm(TARGET_FLAG ~ . - TARGET_AMT, data = insurance.train.df,
                   family = binomial(link = "logit"))

summary(logitModel1)


```


Removing the 4 statistically insignificant features from the model:
1. AGE
2. HOMEKIDS
3. YOJ
4. CAR_AGE

```{r}
logitModel2 <- glm(TARGET_FLAG ~ KIDSDRIV + INCOME + PARENT1 + HOME_VAL + MSTATUS +
                     SEX + EDUCATION + JOB + TRAVTIME + CAR_USE + BLUEBOOK + TIF +
                     CAR_TYPE + RED_CAR + OLDCLAIM +CLM_FREQ + REVOKED + MVR_PTS +
                     URBANICITY
                     , data = insurance.train.df,
                   family = binomial(link = "logit"))

summary(logitModel2)

```

This model looks good. Checking the results and metrics below.
```{r}
predict_logit2 <- predict(logitModel2, newdata = insurance.test.df, type = "response")
predict_logit2_class <- ifelse(predict_logit2 > 0.5, 1, 0)
print(xtabs(~predict_logit2_class + insurance.test.df$TARGET_FLAG))

```

```{r}
roc(insurance.test.df$TARGET_FLAG, predict_logit2, plot = TRUE)

```

Defining the metrics functions.
```{r, echo=FALSE}
Accuracy_func <- function(actual_labels, predicted_labels){
  True_positive_plus_negative <- sum(actual_labels == predicted_labels)
  False_positive_plus_negative <- sum(actual_labels != predicted_labels)
  
  Accuracy <- True_positive_plus_negative / (True_positive_plus_negative + False_positive_plus_negative)
  return(Accuracy)
}

Precision_func <- function(actual_labels, predicted_labels) {
  True_positive <- sum(actual_labels == 1 & predicted_labels == 1)
  False_positive <- sum(actual_labels == 0 & predicted_labels == 1)
  
  Precision_val <- True_positive / (True_positive + False_positive)
  
  return(Precision_val)
}

Sensitivity_func <- function(actual_labels, predicted_labels) {
  True_positive <- sum(actual_labels == 1 & predicted_labels == 1)
  False_negative <- sum(actual_labels == 1 & predicted_labels == 0)
  
  Sensitivity_val <- True_positive / (True_positive + False_negative)
  
  return(Sensitivity_val)
}

Specificity_func <- function(actual_labels, predicted_labels) {
  True_negative <- sum(actual_labels == 0 & predicted_labels == 0)
  False_positive <- sum(actual_labels == 0 & predicted_labels == 1)
  
  Specificity_val <- True_negative / (True_negative + False_positive)
  
  return(Specificity_val)
}

```

```{r}
Accuracy_func(insurance.test.df$TARGET_FLAG, predict_logit2_class)
Precision_func(insurance.test.df$TARGET_FLAG, predict_logit2_class)
Sensitivity_func(insurance.test.df$TARGET_FLAG, predict_logit2_class)
Specificity_func(insurance.test.df$TARGET_FLAG, predict_logit2_class)

```

This looks to be our good model for classifying the data into whether the claim was filed or not by a customer.

### Model to predict the target amount:
We are taking only the rows which have target_flag = 1, as target_flag = 0 means 0 target_amount.

```{r}
insurance_raw_claim <- insurance_raw_imputed_df[insurance_raw_imputed_df$TARGET_FLAG == 1,]
nrow(insurance_raw_claim)

```


```{r}
# Now we will use this dataset to determine / predict the cost

boxplot(insurance_raw_claim$TARGET_AMT)
summary(insurance_raw_claim$TARGET_AMT)

```

Splitting into train and test dataset.
```{r}
n2 <- nrow(insurance_raw_claim)

set.seed(123)
insurance_claim_random_index <- insurance_raw_claim[sample(n2), ]

insurance.claim.train.df <- insurance_claim_random_index[1:as.integer(0.8*n2),]

insurance.claim.test.df <- insurance_claim_random_index[as.integer(0.8*n2 +1):n2, ]

```

Building Linear regression model using all features.
```{r}
linear_model1 <- lm(TARGET_AMT ~ KIDSDRIV + AGE + HOMEKIDS + YOJ + INCOME +
                    PARENT1 +HOME_VAL + MSTATUS + SEX + EDUCATION + JOB + TRAVTIME +
                    CAR_USE + BLUEBOOK + TIF + CAR_TYPE + RED_CAR + OLDCLAIM + CLM_FREQ +
                    REVOKED + MVR_PTS + CAR_AGE + URBANICITY
                    , data = insurance.claim.train.df)

summary(linear_model1)
```

Removing features.
```{r}

linear_model3 <- lm(TARGET_AMT ~ BLUEBOOK + REVOKED + MVR_PTS + CAR_AGE
                    , data = insurance.claim.train.df)

summary(linear_model3)

```


This is our final model.

### Evaluation dataset
```{r}
eval_df <- read.csv("https://raw.githubusercontent.com/deepakmongia/Data621/master/HW-4/Data/insurance-evaluation-data.csv",
                    header = TRUE, row.names = 1)

dim(eval_df)
anyNA(eval_df)

```

### Check missing data
```{r}

eval_missing_df <- data.frame(apply(eval_df, 2, function(x) length(which(x == ''))))
eval_missing_df

eval_na_df1 <- data.frame(apply(eval_df, 2, function(x) length(which(is.na(x)))))
eval_na_df1

```

#### Converting blank data to NA
```{r}
eval_df$INCOME[eval_df$INCOME == ''] <- NA
eval_df$HOME_VAL[eval_df$HOME_VAL == ''] <- NA
eval_df$JOB[eval_df$JOB == ''] <- NA

eval_df$INCOME <- str_remove_all(eval_df$INCOME, "[$,]") %>% as.integer()
eval_df$HOME_VAL <- str_remove_all(eval_df$HOME_VAL, "[$,]") %>% as.integer()
eval_df$BLUEBOOK <- str_remove_all(eval_df$BLUEBOOK, "[$,]") %>% as.integer()
eval_df$OLDCLAIM <- str_remove_all(eval_df$OLDCLAIM, "[$,]") %>% as.integer()


```

#### Imput missing data
```{r}
eval_df1 <- subset(eval_df, select= -c(TARGET_FLAG, TARGET_AMT, JOB))
eval_df2 <- subset(eval_df, select= c(TARGET_FLAG, TARGET_AMT, JOB))

eval_df_imputed <- mice(data = eval_df1, m = 1,
                        method = "pmm", maxit = 5, seed = 500)

eval_df_imputed_df <- mice::complete(eval_df_imputed, 1)

```

```{r}
eval_final_df <- cbind(eval_df_imputed_df, eval_df2)

head(eval_final_df)
dim(eval_final_df)
eval_final_df <- eval_final_df[,c(23,24, 1:10, 25, 11:22)]
```

#### Classifying the evaluate dataset into claim = 0 or 1
```{r}
predict_eval_logit2 <- predict(logitModel2, newdata = eval_final_df, type = "response")
eval_final_df$TARGET_FLAG <- ifelse(predict_eval_logit2 > 0.5, 1, 0)

```

#### Predicting the claim amount, and generating the final evaluated dataset
```{r}
eval_amt_df <- eval_final_df[eval_final_df$TARGET_FLAG == 1, ]

eval_amt_df$TARGET_AMT <- predict(linear_model3, newdata = eval_amt_df)

eval_final_df <- rbind(eval_final_df[eval_final_df$TARGET_FLAG == 0, ], eval_amt_df)

```

